Paper

### 2018-09-05 Paper.Rmd
add Seattle SNP analysis scripts

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = subset(all.patient.info, Set == 'Training')

@@

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

@@

```{r echo=F}
cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')

@@

arm.loc = arrange(rbind(
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

@@

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

@@

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

@@

#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

@@

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

@@

```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
```

```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

@@

## count arms differently than segs??
df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% group_by(Status, Patient) %>% mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = subset(df2, early | late)
el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

@@

ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)

@@

pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

@@

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

@@

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

@@

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
                                          'High'=sum(length(which(Risk == "High"))),
                                          'n.samples'=length(Patient) )

@@

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology))

@@

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),
  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

@@

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

@@

  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

@@

```{r, echo=F, message=F, warning=F}
back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

@@

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
valCNdf$mean.value = valCNdf$value # but not going to use it

@@

m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
                                            'High'=length(which(Risk == 'High'))/length(Patient),
                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))

@@

pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

@@

hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))

@@

qt = cbind(qt, back.v %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))

pred.confidence.v = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P) )

@@

```{r echo=F, eval=T}
dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
rm(dysplasia.df,coefs,plots,labels)

@@

predA = list.files(dirA, 'predictions.Rdata', full.names = T)
predB = list.files(dirB, 'predictions.Rdata', full.names = T)

files = c()
if (length(predA) < 1)
  files = c(list.files(dirA, pattern='raw', full.names=T))
if (length(predB) < 1)
  files = c(files, list.files(dirB, pattern='raw', full.names=T))

if (length(files) > 0) {
  sampleNames = sub('_raw.txt', '', basename(files))
  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))

  for (n in 1:length(files)) {
    print(n)
    print(sampleNames[n])

    rs = read.table(files[n], sep='\t',header=T)
    if (ncol(rs) < 4)
      stop("File needs to contain 4 columns: chr, start, end, sample_value")

    segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
    segM = as.matrix(segs[,-c(1:3)])
    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
    segM = t(segM)
    for (i in 1:ncol(segM)) 
      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])

    arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
    armsM = as.matrix(arms[,-c(1:3)])
    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
    armsM = t(armsM)
    for (i in 1:ncol(armsM)) 
      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])

    nrow(armsM) == nrow(segM)

    cx.score = score.cx(segM,1)

    mergedDf = subtract.arms(segM, armsM)
    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
    dim(mergedDf)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')

    dspred[n,] = c(preds[,1],RR[,1])
    print(dspred[1:n,])
  }
} else {
  load(predB, verbose=T)
  dspredB = dspred

  load(predA, verbose=T)

  dspred = rbind(dspred, dspredB)
  dspred$Type = NA

  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'

}

ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
  theme(legend.position = 'none') + plot.theme

adj = subset(dspred, Type == 'AdjacentBE')
prg = back[,c('Prediction', 'OR', 'Status')]
prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')

@@

maxP = allB %>% group_by(Patient, Status, brks) %>% summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )

@@

second = subset(rules,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

second = subset(rules2,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
subset(sum.patient.data, Patient %in% p) %>% group_by(Patient) %>% summarise(final.endo-initial.endo)

rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )

@@

risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

@@

```{r clonality, echo=F, message=F, warning=F, eval=F}
## Clonal Heterogenity
patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
  length(PID)
)

>>>>

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
head(all.patient.info)

patient.info = all.patient.info

@@

#pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

#pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

@@

```{r echo=F}
cohortSummary = as.data.frame(all.info %>% dplyr::group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')

@@

arm.loc = arrange(rbind(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

@@

# Average cn value
globalCNdf = globalCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

@@

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

@@

#ptendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  #print(unique(pt$Hospital.Research.ID))
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% dplyr::group_by(PID) %>% summarise(max(Pathology))

@@

### TODO: I need to redo the calibration, my confidence intervals seem wrong.

@@

qt = cbind(qt, back %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))

pred.confidence = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

annotation.table = arrange(pred.confidence %>% dplyr::group_by(Risk) %>% dplyr::summarise(
  'n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 
  'min'=min(r1), 'max'=max(r2)), min
)

@@

```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% dplyr::group_by(Risk, min,max) %>% dplyr::summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
```

```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% dplyr::group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

@@

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% dplyr::group_by(Status, Patient) %>% dplyr::mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = subset(df2, early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

@@

#ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
#ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)

@@

pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

@@

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

@@

pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))

@@

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
                                          'High'=sum(length(which(Risk == "High"))),
                                          'n.samples'=length(Patient) )

@@

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology))

@@

endos.m = subset(endos, Status == 'P') %>% dplyr::group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),
  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% dplyr::group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

@@

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% dplyr::group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

@@

  summary.t = summary.t %>% dplyr::group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

@@

```{r, echo=F, message=F, warning=F}
back = back %>% dplyr::group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

@@

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
valCNdf = valCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
valCNdf$mean.value = valCNdf$value # but not going to use it

@@

m = melt(back.v %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
                                            'High'=length(which(Risk == 'High'))/length(Patient),
                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))

@@

pathTotal = back.v %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

@@

hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% dplyr::group_by(Risk) %>% summarize (
  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))

@@

qt = cbind(qt, back.v %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))

pred.confidence.v = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P) )

@@

```{r echo=F, eval=T}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
rm(dysplasia.df,coefs,plots,labels)

@@

dirA = '~/Data/Ellie/Analysis/downsampled'
ds_info = read_excel('~/Data/Ellie/Analysis/downsampled_ids.xlsx')
predA = list.files(dirA, 'predictions.Rdata', full.names = T)
load(predA)

dspred = merge(dspred, ds_info, by.x='row.names', by.y='Illumina ID')

ggplot(dspred, aes(Type, RR, group=Type, fill=Type)) + geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2, aes(color=Status)) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
  plot.theme

adj = subset(dspred, grepl('Adjacent',Type))
prg = back[,c('Prediction', 'OR', 'Status')]
prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')

@@

maxP = allB %>% dplyr::group_by(Patient, Status, brks) %>% dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )

@@

second = subset(rules,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

second = subset(rules2,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
subset(sum.patient.data, Patient %in% p) %>% dplyr::group_by(Patient) %>% dplyr::summarise(final.endo-initial.endo)

rrPath = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )

@@

risksL = subset(bd, Status == 'NP') %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

@@

```{r clonality, echo=F, message=F, warning=F, eval=F}
## Clonal Heterogenity
patient.info %>% dplyr::group_by (Hospital.Research.ID, Patient, PID) %>% dplyr::summarise(
  length(PID)
)
